---
title: "Plotting roads example"
output: html_notebook
---

Attempting to plot roads on a map and colour the roads according to data. The idea is to colour roads according to their congestion.

Road data is obtained from open street map. A website converts road data into shapefiles.
http://download.geofabrik.de/index.html


```{r, message = F, warning = F}
library(tidyverse)
library(leaflet)
library(rgdal)
```

Download the data for Cambridgeshire as an example

```{r, message = F}
download.file('http://download.geofabrik.de/europe/great-britain/england/cambridgeshire-latest-free.shp.zip', 'cambridgeshire.zip' )
unzip('cambridgeshire.zip', exdir = 'data')
```

Laod the shapefile corresponding to roads

```{r}
shapefile <- readOGR('data/gis.osm_roads_free_1.shp')
summary(shapefile@data)
```

Filter down only to trunk roads and motorways
Use rmapshaper to simplify and reduce the file size
```{r}
f = shapefile$fclass == 'trunk' | shapefile$fclass == 'motorway'
main_roads <- shapefile[f, ]
object.size(main_roads)
main_roads_simple  <- rmapshaper::ms_simplify(main_roads)
object.size(main_roads_simple)
```

rmapshaper only had a little effect. Maybe use mapshaper.org to simplify the lines more

Plot on a map

```{r}
pal <- colorNumeric('BuGn', domain = c(0, 125))
leaflet(main_roads) %>%
  addTiles() %>%
  addPolylines(opacity = 1, color = pal(main_roads$maxspeed)) %>%
  addLegend(position = 'bottomleft', pal = pal, values = seq(0, 125, by = 25), opacity = 1,
            title = 'Speed kph')
```

